The Impact of Accretion Disk Winds on the Optical Spectra of Cataclysmic Variables

### 2014-12-03 draft.tex
added some more CK changes

\begin{abstract}
Many high-state non-magnetic cataclysmic variables (CVs) exhibit
blue-shifted absorption or P-Cygni profiles in ultraviolet
(UV) resonance lines, such as C~{\sc iv}~1550\AA. These features
imply the existence of powerful accretion disk winds in
CVs, and considerable effort has gone into modelling UV line formation
in such outflows. By contrast, much less attention has been paid to
the potential impact of these disk winds on the {\em optical} spectra
of CVs. Here, we use an improved version of our Monte Carlo ionization
and radiative transfer code to investigate whether disk wind models that
produce realistic UV line profiles are also likely to generate
observationally significant recombination line and continuum emission
in the optical waveband. In particular, we test whether outflows may 
be responsible for the single-peaked emission line profiles often seen
in high-state CVs and for the weakness of the Balmer absorption edge
(relative to simple models of optically thick accretion disks).
We find that a standard disk wind model that is successful in
reproducing the UV spectra of CVs does leave a noticeable imprint on
the optical spectrum as well, particularly for systems viewed at high
inclination. The strongest optical wind-formed recombinations lines are \ha\ and
\heiiopt\, although the outflow also produces emission in the Ly-$\alpha$
and \heiiuv\ UV lines. However, all of the optical lines produced by
this default wind model are double-peaked, and the effect of the wind
on the Balmer jump is modest. We carry out a limited exploration of
the disk wind parameter space, focusing on those parameters that affect the
density near the base of the wind. This shows that higher-density
outflows can produce both single-peaked emission lines and a
recombination continuum sufficient to fill in the Balmer
jump, although such models also tend to predict stronger-than-observed 
collisionally excited UV emission lines. Our main conclusion is that
disk wind emission is likely to have a significant impact on the
optical spectra of high-state CVs. 
\end{abstract}

@@

Much less is known about the effect of these outflows on the optical
spectra of high-state CVs. These spectra are typically characterized
by H and He emissions superposed on a blue continuum. In many
cases, and particularly in the SW~Sex subclass of NLs
\citep{HSK86,DR95}, these lines are single-peaked. This is contrary to
theoretical expectations for lines formed in accretion disks, which
are predicted to be double-peaked \citep{smak1981, hornemarsh1986}. 
{\em Low-state} CVs (dwarf novae in quiescence) do, in fact,
exhibit such double-peaked lines \citep{marshhorne1990}(MOREREF). 
{\bf JM: Other things to consider; transition of line profiles
from quiescence to outburst in DNe, emission lines in absorption cores
trends of lines and continuum with inclination, flat
Balmer decrement}.

Murray \& Chiang (1996) have shown that the presence of disk winds may
offer a natural explanation for the single-peaked optical emission lines in
high-state CVs, since they can strongly affect the radiative transfer
of line photons. Strong support for a significant wind contribution to the
optical emission lines comes from observations of eclipsing
systems. There, the single-peaked lines are often only weakly
eclipsed, and a significant fraction of the line flux remains visible
even near mid-eclipse (REF). This points to line formation in a spatially
extended region, such as a disk wind (see figure~\ref{novalikes}).
%%(Honeycutt, Schlegel, \& Kaitchuck 1986; Dhillon \& Rutten 1995). 
Further evidence for a wind contribution to the optical lines comes
from isolated observations of P-Cygni-like line profiles even in optical
lines, such as \ha\ and He \textsc{i} $\lambda5876$ \citep{RN98}.
{\bf JM: need to mention this as pointed out by CK.}

@@

Prior to SDL05, the relative ionization fractions for all atomic
species were estimated via the modified Saha equation (Mazzali \&
Lucy 1993)  
\begin{equation}
\frac{n_{j+1} n_e}{n_j} = W [\xi + W(1-\xi)]
\left(\frac{T_e}{T_R}\right)^{1/2}
\left(\frac{n_{j+1}n_e}{n_j}\right)^*_{T_R}, \label{ionization}
\end{equation}
Here, the `starred' term on the right represents the Saha (LTE)
abundances, in accordance with the notation of Mihalas (1982).  
$W$ is an effective dilution factor, $\xi$ is the
fraction of recombinations going directly to the ground state, and
$T_R$ and $T_e$ are the radiation and electron temperature
respectively. This simple ionization scheme produces reasonable
results when the photoionizing SED can be approximated by a dilute
blackbody. This is the case for high-state CVs. (As noted above, an
improved, but more complex treatment of ionization that is appropriate
for more complex SEDs is described in H13.)

Similarly, the relative excitation fractions within each ionization
stage of a given species were estimated via a modified Boltzmann
equation,
\begin{equation}
\frac{n_{jk}}{n_j} = W g_k \exp{-E_k/kT_R} / z_j(T_R),
\end{equation}
where $n_{jk}$ is the population of level $k$ in ionic stage $j$,
$E_k$ is the energy difference between level $k$ and the ground state,
and $z_j(T_R)$ is the partition function of ionic stage $j$.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/fig_caseb_tardis.eps}
\caption{
{\sl Top Panel:} Balmer decrements calculated with \py compared with analytic calculations
by Seaton (1959), showing good agreement when calculating level emissivities in Hydrogen
in the Case B limit.
(see Osterbrock 1989 for a discussion of this commonly used approximation).
{\sl Bottom Panel:}  a comparison of Helium I level populations (the most complex ion we currently 
treat as a macro-atom) between \py~and \tar~models. Considering the 
two codes use different atomic data and \textsc{Tardis,} unlike \textsc{Python,} currently has 
a complete treatment of collisions between radiatively forbidden transitions, the factor of 
$<2$ agreement is encouraging. 
{\bf CK: WHAT ARE THE PARTICULAR SETS OF PARAMETERS BEING ADOPTED HERE?}
}
\label{tests}
\end{figure}
\nocite{osterbrock}
\nocite{seaton1959}

@@

The density at position $(r,z)$ in the wind, $\rho(r,z)$, is
calculated from the mass continuity equation, yielding
\begin{equation}
\rho(r,z) = \frac{r}{r_0} \frac{dr}{dr_0} \frac{\phi(r_0)}{v_z(r,z)}.
\label{density}
\end{equation}
Here, $\phi(r_O)$ is the mass-loss rate per unit area at $r_0$, 
and $v_z$ is the vertical velocity component. We take $\phi(r_0)$ to be
uniform across the wind-launching area of the disk and normalize it by
matching its integral to the user-specified total mass-loss rate,
$\dot{M}_{wind}$. 
{\bf CK: I'm still not sure I understand the formula
  SV93 use to define $\phi$, and hence that this description is
  actually strictly correct.... Not sure what to do about that
  though...}

@@

It is possible to include radiation from a boundary layer (BL) between
the disk and the WD. In \py, the BL is described as
a blackbody with a user-specified effective temperature and
luminosity. In the models presented here, we have followed LK02 in setting
the BL luminosity to zero. However, we have confirmed that the addition of a
BL with $L_{BL} = 0.5 L_{acc}$ and temperatures in the range $80{\rm
kK} \leq T_{BL} \leq 200{\rm kK}$ would not change any of our main
conclusions. 
{\bf CK: Is this true? Do we want to say anything else, e.g. that we'll
     talk about this in a future paper?}

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/sed_figure.eps}
\caption{
The spectral energy distribution (SED) of the blackbody accretion
disk used in the ionization calculation (black line), 
with two boundary layer blackbody spectra also plotted.
The positions of the Hydrogen and Helium ionization edges 
are marked with dashed vertical lines.
{\bf CK: this will need to be updated -- we want to show (just) the disk and
  WD spectra...}
 }
\label{sed}
\end{figure}

@@

Figure~\ref{spec} shows the corresponding optical spectra produced for
the benchmark model, and these do exhibit some emission lines
associated with H and He. Since the emissivity of these recombination 
features scales as $n_e^2$, they are formed almost entirely in the 
dense base of the wind, just above the accretion disk. Here, the
velocity field of the wind is still dominated by rotation, rather than
outflow, which accounts for the double-peaked shape of the lines. In
principle, lines formed in this region can still be single peaked,
since the existence of a poloidal velocity {\em gradient} changes the
local escape probabilities (Murray \& Chiang 1996). However, as
demonstrated explicitly in Section~\ref{BLA}, the line opacity in our
benchmark model is not high enough for this radiative transfer effect
to dominate the line shapes.

@@

The results of Section~\ref{BLA} show that wind models that reproduce
the UV spectral features of CVs are likely to leave a clear imprint on
the optical spectrum. It is therefore important to understand the
sensitivity of optical (and UV) wind-formed features to the basic
properties of the outflow. In particular, there are two key questions
we would like to address: (i) under what conditions can disk winds
produce {\em single-peaked} Balmer emission lines (Murray \& Chiang
1996); (ii) can the wind-formed recombination continuum completely
fill in the disk's Balmer absorption edge for reasonable outflow
parameters?

@@

Two other features are worth noting in this part of the
spectrum. First, the collisionally excited Ca~{\sc ii} H\&K 
emission lines at 3969\AA 
and (especially) 3934\AA become quite prominent in our densest
models. Second, {\em all} of our models predict a detectable
He~\textsc{ii} recombination line at 3202\AA. This is the Helium
equivalent of Paschen$\beta$ and should be expected in all systems that
feature a strong He~\textsc{ii}~4686\AA line (the Helium
equivalent of Paschen$\alpha$). This line is unfamiliar observationally,
but only because it lies bluewards of the atmospheric cut-off, but
also redwards of most ultraviolet spectra. 
{\bf JM: should we reference something for this paper here}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_alpha.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_beta.eps}}   
}
\caption{
Line profiles for \ha\ (left) and \hb\ (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$.
The benchmark model is shown in the top left panel, and the improved optical
model described in section 6 is highlighted in bold. 
The parameter that has been altered from the benchmark model 
is marked in each column and row. 
Models towards the bottom right
are have more slowly accelerating winds and denser wind bases. }
\label{halpha}
\end{figure} %fullpage

@@

The dependence of the ultraviolet C~{\sc iv} and He~{\sc ii} lines on
the same kinematic wind parameters is illustrated in Figures~\ref{BLA}
and~\ref{BLA}. As before, we focus on the predicted behaviour at $i =
80^\circ$. In our benchmark model, the C~{\sc iv} emission
line seen at high inclinations is almost entirely produced via
resonance scattering. The angle-averaged flux integrated over the
entire line profile -- accounting correctly for the blue-shifted
absorption component seen at low inclinations -- would be
approximately zero. However, as we increase the density and extent of
the slow-moving part of the outflow near the disk plane, collisionally
excited emission starts to contribute significantly to the C~{\sc iv}
line profiles. This explains the much stronger line-to-continuum
contrast seen in the $\alpha = 1.5$, $R_v = 2\times 10^{11}$~cm model,
for example. For any given inclination, the trend is not necessarily
monotonic, however, because the escape probabilities from the
line-forming region can be highly anisotropic.  
{\bf CK: I *think* this is the basic reason, but (a) can you check
  this, James; (b) is it enough to just point this out?; (c) do we
  understand why CIV is single-peaked but HeII, for example, is double-peaked?}

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_1640.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_4686.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
Line profiles for \heiiuv\ (left) and \heiiopt\  (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$. 
The panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{heiifig}
\end{figure}

>>>>

\begin{abstract}
Many high-state non-magnetic cataclysmic variables (CVs) exhibit
blue-shifted absorption or P-Cygni profiles in ultraviolet
(UV) resonance lines, such as C~{\sc iv}~1550\AA. These features
imply the existence of powerful accretion disk winds in
CVs, and considerable effort has gone into modelling UV line formation
in such outflows. By contrast, much less attention has been paid to
the potential impact of these disk winds on the {\em optical} spectra
of CVs. Here, we use an improved version of our Monte Carlo ionization
and radiative transfer code to investigate whether disk wind models that
produce realistic UV line profiles are also likely to generate
observationally significant recombination line and continuum emission
in the optical waveband. In particular, we test whether outflows may 
be responsible for the single-peaked emission line profiles often seen
in high-state CVs and for the weakness of the Balmer absorption edge
(relative to simple models of optically thick accretion disks).
We find that a standard disk wind model that is successful in
reproducing the UV spectra of CVs does leave a noticeable imprint on
the optical spectrum as well, particularly for systems viewed at high
inclination. The strongest optical wind-formed recombinations lines are \ha\ and
\heiiopt\, although the outflow also produces emission in the Ly-$\alpha$
and \heiiuv\ UV lines. However, all of the optical lines produced by
this default wind model are double-peaked, and the effect of the wind
on the Balmer jump is modest. We carry out a limited exploration of
the disk wind parameter space, focusing on those parameters that affect the
density near the base of the wind. This shows that higher-density
outflows can produce both single-peaked emission lines and a
recombination continuum sufficient to fill in the Balmer
jump, although such models also tend to predict stronger-than-observed 
collisionally excited UV emission lines. Our main conclusion is that
disk wind emission is likely to have a significant impact on the
optical spectra of high-state CVs. 
{\bf NOTE JM: This abstract is a few lines longer than I would like- 
I think something punchier is required.}
\end{abstract}

@@

Much less is known about the effect of these outflows on the optical
spectra of high-state CVs. These spectra are typically characterized
by H and He emissions superposed on a blue continuum. In many
cases, and particularly in the SW~Sex subclass of NLs
\citep{HSK86,DR95}, these lines are single-peaked. This is contrary to
theoretical expectations for lines formed in accretion disks, which
are predicted to be double-peaked \citep{smak1981, hornemarsh1986}. 
{\em Low-state} CVs (dwarf novae in quiescence) can, in fact,
exhibit such double-peaked lines \citep{marshhorne1990}. 
{\bf NOTEJM: I've included a reference to a paper on IP Peg here
which shows the transition from double peaked lines in 
quiescence and a contrast to outburst spectra. I'm not sure
of general trends in DNe. A lot seem to have weak emission cores
embedded in absorption.}

Murray \& Chiang (1996; hereafter MC96)\nocite{MC96} 
have shown that the presence of disk winds may
offer a natural explanation for the single-peaked optical emission lines in
high-state CVs, since they can strongly affect the radiative transfer
of line photons. Strong support for a significant wind contribution to the
optical emission lines comes from observations of eclipsing
systems. There, the single-peaked lines are often only weakly
eclipsed, and a significant fraction of the line flux remains visible
even near mid-eclipse (REF). This points to line formation in a spatially
extended region, such as a disk wind (see figure~\ref{novalikes}).
%%(Honeycutt, Schlegel, \& Kaitchuck 1986; Dhillon \& Rutten 1995). 
Further evidence for a wind contribution to the optical lines comes
from isolated observations of P-Cygni-like line profiles even in optical
lines, such as \ha\ and He \textsc{i} $\lambda5876$ \citep{RN98}.
{\bf JM: need to mention this as pointed out by CK.}

@@

Prior to SDL05, the relative ionization fractions for all atomic
species were estimated via the modified Saha equation (Mazzali \&
Lucy 1993)  
\begin{equation}
\frac{n_{j+1} n_e}{n_j} = W [\xi + W(1-\xi)]
\left(\frac{T_e}{T_R}\right)^{1/2}
\left(\frac{n_{j+1}n_e}{n_j}\right)^*_{T_R}, \label{ionization}
\end{equation}
Here, the `starred' term on the right represents abundances computed with
the Saha equation at temperature $T_R$, but using actual partition functions
from the dilute blackbody approximation. 
$W$ is an effective dilution factor, $\xi$ is the
fraction of recombinations going directly to the ground state, and
$T_R$ and $T_e$ are the radiation and electron temperature
respectively. This simple ionization scheme produces reasonable
results when the photoionizing SED can be approximated by a dilute
blackbody. This is the case for high-state CVs. (As noted above, an
improved, but more complex treatment of ionization that is appropriate
for more complex SEDs is described in H13.)

Similarly, the relative excitation fractions within each ionization
stage of a given species were estimated via a modified Boltzmann
equation,
\begin{equation}
\frac{n_{jk}}{n_j} = \frac{W g_k}{z_j(T_R)} \exp(-E_k/kT_R),
\end{equation}
where $n_{jk}$ is the population of level $k$ in ionic stage $j$,
$E_k$ is the energy difference between level $k$ and the ground state,
and $z_j(T_R)$ is the partition function of ionic stage $j$.

@@

\begin{figure}
\centering
\includegraphics[width=0.5\textwidth]{figures/fig_caseb_tardis.eps}
\caption{
{\sl Top Panel:} Balmer decrements calculated with \py compared with analytic calculations
by Seaton (1959), showing good agreement when calculating level emissivities in Hydrogen
in the Case B limit.
(see Osterbrock 1989 for a discussion of this commonly used approximation).
{\sl Bottom Panel:}  a comparison of Helium I level populations (the most complex ion we currently 
treat as a macro-atom) between \py~and \tar~models. Considering the 
two codes use different atomic data and \textsc{Tardis,} unlike \textsc{Python,} currently has 
a complete treatment of collisions between radiatively forbidden transitions, the factor of 
$<2$ agreement is encouraging. 
{\bf NOTE CK: WHAT ARE THE PARTICULAR SETS OF PARAMETERS BEING ADOPTED HERE?}
}
\label{tests}
\end{figure}
\nocite{osterbrock}
\nocite{seaton1959}

@@

The density at position $(r,z)$ in the wind, $\rho(r,z)$, is
calculated from the mass continuity equation, yielding
\begin{equation}
\rho(r,z) = \frac{r}{r_0} \frac{dr}{dr_0} \frac{\phi(r_0)}{v_z(r,z)}.
\label{density}
\end{equation}
Here, $\phi(r_O)$ is the mass-loss rate per unit area at $r_0$, 
and $v_z$ is the vertical velocity component. We take $\phi(r_0)$ to be
uniform across the wind-launching area of the disk and normalize it by
matching its integral to the user-specified total mass-loss rate,
$\dot{M}_{wind}$. 
{\bf NOTE CK: I'm still not sure I understand the formula
  SV93 use to define $\phi$, and hence that this description is
  actually strictly correct.... Not sure what to do about that
  though...}

@@

It is possible to include radiation from a boundary layer (BL) between
the disk and the WD. In \py, the BL is described as
a blackbody with a user-specified effective temperature and
luminosity. In the models presented here, we have followed LK02 in setting
the BL luminosity to zero. However, we have confirmed that the addition of a
BL with $L_{BL} = 0.5 L_{acc}$ and temperatures in the range $80{\rm
kK} \leq T_{BL} \leq 200{\rm kK}$ would not change any of our main
conclusions. 
{\bf NOTE CK: Is this true? Do we want to say anything else, e.g. that we'll
talk about this in a future paper? JM: This is true. Anything up to 200kK
still has optical lines. The He II behaviour is not monotnic. CIV drops
gradually and gets a little overionized- but this is a separate topic really.
The issue of clumping and BL and their affect on ionization could be a 
future paper.}

@@

\begin{figure}
\centering
\includegraphics[width=0.45\textwidth]{figures/sed_figure.eps}
\caption{
The spectral energy distribution (SED) of the blackbody accretion
disk used in the ionization calculation (black line), 
with two boundary layer blackbody spectra also plotted.
The positions of the Hydrogen and Helium ionization edges 
are marked with dashed vertical lines.
{\bf JM: do we want to show the stellar atmosphere spectrum here?
we know it looks a bit crap around the He II edge, so perhaps not...
 }
\label{sed}
\end{figure}

@@

Figure~\ref{spec} shows the corresponding optical spectra produced for
the benchmark model, and these do exhibit some emission lines
associated with H and He. Since the emissivity of these recombination 
features scales as $n_e^2$, they are formed almost entirely in the 
dense base of the wind, just above the accretion disk. Here, the
velocity field of the wind is still dominated by rotation, rather than
outflow, which accounts for the double-peaked shape of the lines. In
principle, lines formed in this region can still be single peaked,
since the existence of a poloidal velocity {\em gradient} changes the
local escape probabilities (MC96). However, as
demonstrated explicitly in Section~\ref{BLA}, the line opacity in our
benchmark model is not high enough for this radiative transfer effect
to dominate the line shapes.

@@

The results of Section~\ref{BLA} show that wind models that reproduce
the UV spectral features of CVs are likely to leave a clear imprint on
the optical spectrum. It is therefore important to understand the
sensitivity of optical (and UV) wind-formed features to the basic
properties of the outflow. In particular, there are two key questions
we would like to address: (i) under what conditions can disk winds
produce {\em single-peaked} Balmer emission lines (MC96); (ii) can the wind-formed recombination continuum completely
fill in the disk's Balmer absorption edge for reasonable outflow
parameters?

@@

Two other features are worth noting in this part of the
spectrum. First, the collisionally excited Ca~{\sc ii} H\&K 
emission lines at 3969\AA 
and (especially) 3934\AA become quite prominent in our densest
models. Second, {\em all} of our models predict a detectable
He~\textsc{ii} recombination line at 3202\AA. This is the Helium
equivalent of Paschen~$\beta$ and should be expected in all systems that
feature a strong He~\textsc{ii}~4686\AA\ line (the Helium
equivalent of Paschen~$\alpha$). This line is unfamiliar observationally,
but only because it lies bluewards of the atmospheric cut-off, but
also redwards of most ultraviolet spectra. 
{\bf JM: should we reference something for this paper here}

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_alpha.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_4686.eps}}   
}
\caption{
Line profiles plotted in veliocity space 
for \ha\ (left) and \heiiopt\ (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$.
The benchmark model is shown in the top left panel, and the improved optical
model described in section 6 is highlighted in bold. 
The parameter that has been altered from the benchmark model 
is marked in each column and row. Models towards the bottom right
are have more slowly accelerating winds and denser wind bases. }
\label{halpha}
\end{figure} %fullpage

\begin{figure} %fullpage
\centering
\includegraphics[width=0.8\textwidth]{figures/3by3_grid_balmer.eps}  
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
Line profiles for \civ\ (left) and \heiiuv\  (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$. 
The panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{uvlines}
\end{figure} %fullpage

@@

The dependence of the ultraviolet C~{\sc iv} and He~{\sc ii} lines on
the same kinematic wind parameters is illustrated in Figures~\ref{BLA}
and~\ref{BLA}. As before, we focus on the predicted behaviour at $i =
80^\circ$. In our benchmark model, the C~{\sc iv} emission
line seen at high inclinations is almost entirely produced via
resonance scattering. The angle-averaged flux integrated over the
entire line profile - accounting correctly for the blue-shifted
absorption component seen at low inclinations - would be
approximately zero. However, as we increase the density and extent of
the slow-moving part of the outflow near the disk plane, collisionally
excited emission starts to contribute significantly to the C~{\sc iv}
line profiles. This explains the much stronger line-to-continuum
contrast seen in the $\alpha = 1.5$, $R_v = 2\times 10^{11}$~cm model,
for example. For any given inclination, the trend is not necessarily
monotonic, however, because the escape probabilities from the
line-forming region can be highly anisotropic.  
{\bf CK: I *think* this is the basic reason, but (a) can you check
  this, James; (b) is it enough to just point this out?; (c) do we
  understand why CIV is single-peaked but HeII, for example, is double-peaked?}

@@

\begin{figure} %fullpage
\mbox{
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_uv.eps}}
\quad
\subfigure{\includegraphics[width=0.5\textwidth]{figures/3by3_grid_1640.eps}}   
}
%%\includegraphics[width=0.5\textwidth]{figures/test.eps}
\caption{
Line profiles for \civ\ (left) and \heiiuv\  (right) with varying kinematic 
properties, computed for an inclination of $80^\circ$. 
The panels are organised in the same way as figure~\ref{halpha}.
% The panels are ordered in the same way as figure~\ref{halpha}. In general, increasing densities
% lead to a Balmer jump in emission, although as in the case of H$\alpha$ there is a transition
% when the wind becomes a neutral absorbing column.
}
\label{uvlines}
\end{figure}
